% !TEX root =  fission.tex

\bibitem{Verbeke 2016} J.M. Verbeke, J. Randrup, R. Vogt, ``Fission Reaction Event Yield Algorithm FREYA 2.0 User Manual,'' LLNL-TR-XXXXXX , Lawrence Livermore National Laboratory, Livermore, California (2016).

\bibitem{Terrell 1957} J. Terrell, "Distributions of Fission Neutron
Numbers", \textit{Phys. Rev.} \textbf{108}, 783 (1957).

\bibitem{Zucker and Holden 1986} M.S. Zucker, N.E. Holden, "Energy
Dependence of Neutron Multiplicity P$_{\nu}$ in Fast-Neutron-Induced
Fission for $^{235,238}U$ and $^{239}Pu$," BNL-38491 (1986).

\bibitem{Cullen 2006} D.E. Cullen, "Sampling the Number of Neutrons
Emitted per Fission," UCRL-TR-222526, Lawrence Livermore National
Laboratory (2006).  

\bibitem{Gwin 1984} R. Gwin, R.R. Spencer,
R.W. Ingle, "Measurements of the Energy Dependence of Prompt Neutron
Emission from $^{233}U$, $^{235}U$, $^{239}Pu$, and $^{241}Pu$ for
$E_n=0.005$ to 10 eV Relative to Emission from Spontaneous Fission of
$^{252}$Cf," \textit{Nucl. Sci. Eng.}, \textbf{87}, 381 (1984).

\bibitem{Valentine 1996} T.E. Valentine, J.T. Mihalczo, "MCNP-DSP: A
Neutron and Gamma Ray Monte Carlo Calculation of Source-Driven
Noise-Measured Parameters ," \textit{Ann. of Nucl. Eng.}, \textbf{23},
16, p. 1271 (1996).

\bibitem{Valentine 2000} T.E. Valentine, "MCNP-DSP Users Manual,"
ORNL/TM-13334, R2, Oak Ridge National Laboratory (2000).

\bibitem{Frehaut 1988} J. Frehaut, "Neutron Multiplicity Distribution
in Fast Neutron-Induced Fission," \textit{Proc. of IAEA Consultant's
Meeting on Physics of Neutron Emission in Fission,}, Mito, Japan
(1988).  

\bibitem{Holden and Zucker BNL} N.E. Holden,
M.S. Zucker, "A Reevaluation of the Average Prompt Neutron Emission
Multiplicity ($\bar{\nu}$) Values from Fission of Uranium and
Transuranium Nuclides," BNL-NCS-35513, Brookhaven National
Laboratory).  

\bibitem{Santi 2005} P. Santi, D.H. Beddingfield, D.R. Mayo, "Revised
prompt neutron emission multiplicity distributions for $^{236,238}$Pu,"
 \textit{Nucl. Phys. A} \textbf{756}, 325-332 (2005).

\bibitem{BNL-36467} BNL-36467.

\bibitem{Dakavoski 1973} Dakavoski, \textit{Sov.Atom.Erg.} \textbf{17},
360, BNL-36467, (1973).

\bibitem{Hoffman 1980} Hoffman, \textit{Phys.Rev.C.} \textbf{21},
637, BNL-36467, (1980).

\bibitem{Lazarev 1974} Lazarev, \textit{Phys.Lett.} \textbf{52B},
321, BNL-36467, (1974).

\bibitem{Spencer 1982} R.R. Spencer, R. Gwin, R.W. Ingle, "A
measurement of the Average Number of Prompt Neutrons from Spontaneous
Fission of Californium-252," \textit{Nucl. Sci. Eng.} \textbf{80}, 603
(1982).  

\bibitem{Boldeman 1985} J.W. Boldeman, M.G. Hines, "Prompt
Neutron Emission Probabilities Following Spontaneous and Thermal
Neutron Fission," \textit{Nucl. Sci. Eng.}, \textbf{91}, 114 (1985).

\bibitem{Ensslin 1998} N. Ensslin, W.C. Harker, M.S. Krick,
D.G. Langner, M.M. Pickrell, J.E. Stewart, "Application Guide to
Neutron Multiplicity Counting," LA-13422-M, Los Alamos National
Laboratory (1998).  

\bibitem{ENDL 1975} R.J. Howerton, et al, "The LLL
Evaluated Nuclear Data Library (ENDL): Evaluation Techniques, Reaction
Index, and Description of Individual Evaluations," UCRL-50400, V. 15,
Part A, Lawrence Livermore National Laboratory (1975).

\bibitem{TART 2003} D.E. Cullen, "TART 2002: A Couple
Neutron-Photon 3-D, Combinatorial Geometry, Time Dependent Monte-Carlo
Transport Code," UCRL-ID-126455, Rev. 4, Lawrence Livermore National
Laboratory (2003).  

\bibitem{Cullen 2004} D.E. Cullen, "Sampling ENDL Watt Fission
Spectra," UCRL-TR-203251, Lawrence Livermore National Laboratory
(2004).  

\bibitem{Everett 1983} C.J. Everett, E.D. Cashwell, "A Third
Monte Carlo Sampler," LA-9721-MS, Los Alamos National Laboratory
(1983).  


\bibitem{Mannhart 1987} W. Mannhart, "Evaluation
of the Cf-252 Fission Neutron Spectrum Between 0 MeV and 20 MeV,"
\textit{Proc. Advisory Group Mtg. Neutron Sources}, Leningrad, USSR,
1986 (IAEA-TECDOC-410), Vienna (1987).  

\bibitem{Madland 1984}
D.G. Madland, J.R. Nix, "Prompt Fission Neutron Spectra and Average
Prompt Neutron Multiplicities,"NEANDC Specialist's Meeting on Yields
and Decay Data of Fission Products, Brookhaven National Laboratory,
BNL 51778 (1984).  

\bibitem{Froehner 1990} F.H. Fr\"{o}hner,
"Evaluation of $^{252}$Cf Prompt Fission Neutron Data from 0 to 20 MeV
by Watt Spectrum Fit," \textit{Nucl. Sci. Eng.} \textbf{106}, 345
(1990).  

\bibitem{Beck 2007} B. Beck, D.A. Brown, F. Daffin, J. Hedstrom, 
R. Vogt, "Implementation of Energy-Dependent Q Values for Fission,"
UCRL-TR-234617, Lawrence Livermore National Laboratory (2007).  

\bibitem{Madland 1982} D. Madland, J.R. Nix, \textit{Nucl. Sci. and
Eng.} \textbf{81}, p.213 (1982).

\bibitem{Vogt 2008} R. Vogt, "Energy-Dependent Fission Q Values 
Generalized for All Actinides," LLNL-TR-407620, Lawrence Livermore 
National Laboratory (2008).  

\bibitem{Brunson 1982} G.S. Brunson, Jr., "Multiplicity and
Correlated Energy of Gamma Rays Emitted in the Spontaneous Fission of
Californium-252," Ph.D. Thesis, University of Utah (1982).

\bibitem{Valentine 2001} T.E. Valentine, "Evaluation of Prompt Fission
Gamma Rays for Use in Simulating Nuclear Safeguard Measurements,"
\textit{Ann. Nucl. Eng.}, \textbf{28}, 191 (2001).  

\bibitem{Wagemans
1991} C. Wagemans, "The Nuclear Fission Process," \textit{CRC Press,
Inc.,} Boca Raton, Florida (1991).  

\bibitem{Maienschein 1958}
F.C. Maienschein, R.W. Peelle, T.A. Love, \textit{Neutron
Phys. Ann. Prog. Rep. for Sept. 1, 1958}, ORNL-2609, Oak Ridge
National Laboratory (1958).  

\bibitem{Goldstein 1959} "Fundamental
Aspects of Reactor Shielding," Addison-Wesley Publishing Company,
Inc. Reading, Massachussetts (1959).

\bibitem{Oblozinsky 1998} "Summary Report of the $2^{nd}$ Research Coordination 
Meeting on Compilation and Evaluation of Photonuclear Data for 
Applications," International Atomic Energy Agency Report INDC(NDS)-384, 
Vienna, Austria (1998).

%\bibitem{McLane 1997} V. McLane, C.L. Dunford and P.F. Rose, "ENDF-102 Data 
%Formats and Procedures for the Evaluated Nuclear Data File ENDF-6," Report 
%BNL-NCS-44945 Rev. 2/97, Brookhaven National Laboratory, Upton, New York (1997).

\bibitem{Chadwick 1999} M.B. Chadwick et al., "Cross-Section Evaluations to 
150 MeV for Accelerator-Driven Systems and Implementation in MCNPX," 
\textit{Nuclear Science and Engineering,} \textbf{131,} pp. 293-328 (1999).

\bibitem{IAEAphoto} P. Oblo{\v z}insk{\' y}, ed. ``Handbook of photonuclear data for applications: Cross sections and spectra,'' International Atomic Energy Association report IAEA-TECDOC-1178, Vienna, Austria (2000).

\bibitem{ENDFB7} M.B. Chadwick, P. Oblo{\v z}insk{\' y}, M. Herman, N.M. Greene, R.D. McKnight, D.L. Smith, P.G. Young, R.E. MacFarlane, G.M. Hale, S.C. Frankle, A.C. Kahler, T. Kawano, R.C. Little, D.G. Madland, P. Moller, R.D. Mosteller, P.R. Page, P. Talou, H. Trellue, M.C. White, W.B. Wilson, R. Arcilla, C.L.Dunford, S.F. Mughabghab, B. Pritychenko, D. Rochman, A.A. Sonzogni, C.R. Lubitz, T.H. Trumbull, J.P. Weinman, D.A. Brown, D.E. Cullen, D.P. Heinrichs, D.P. McNabb, H. Derrien,
M.E. Dunn, N.M. Larson, L.C. Leal, A.D. Carlson, R.C. Block, J.B. Briggs, E.T. Cheng, H.C. Huria, M.L. Zerkle, K.S. Kozier, A. Courcelle, V. Pronyaev and S.C. van der Marck, ``{ENDF/B-VII.0}: Next Generation Evaluated Nuclear Data Library for Nuclear Science and Technology,''
{\em Nuclear Data Sheets}, {\textbf 107}, pp.~2931-3060 (2006), ISSN 0090-3752, DOI: 10.1016/j.nds.2006.11.001,
(\url{http://www.sciencedirect.com/science/article/B6WNV-4MGDW8W-1/2/5040b3d5640d1c334634e9f83c754893}).

\bibitem{Firestone 1996} R.B. Firestone, V.S. Shirley, C.M. Baglin, S.Y.F. Chu, J. Zipkin, "The 8th edition of the Table of Isotopes," Wiley \& Sons, Inc. (1996).

\bibitem{Wapstra 2003} A.H. Wapstra, G. Audi, C. Thibault, "The AME2003 atomic mass evaluation (I). Evaluation of input data, adjustment procedures," {\em Nuclear physics} {\textbf A729,} 129 (2003).

\bibitem{Audi 2003} G. Audi, A.H. Wapstra, C. Thibault, "The AME2003 atomic mass evaluation (II). Tables, graphs, and references," {\em Nuclear physics} {\textbf A729,} 337 (2003).
