% Geant4 Physics Manual Entry
%
% Use 'make' to run latex and generate the pdf file
%
% Doug Wright

\documentclass[11pt]{article}
\usepackage{h4}

\newcommand{\notgeant}[1]{}% exclude some things from the geant manual

%\ucrl{UCRL-AR-228518}

\begin{document}

\section{Fission with improved multiplicity sampling}

As an alternative to the fission model described in the previous
section there is a modified model that produces more accurate
multiplicity distributions for the emitted neutrons and gamma rays
from spontaneous and neutron-induced fission. This was motivated by
detailed statistical studies of fission chains in multiplying
media. This model is data-driven and incorporates all available
multiplicity measurements found in the literature. Empirical models
are employed whenever multiplicity data are not available.
Essentially no data are available for the correlations between the
neutrons and gammas, so this model samples these distributions
independently. By default, this model effectively scales the
multiplicity data to match the average multiplicity value
($\bar{\nu}$) found in the GEANT4 evaluated data library. Therefore,
only isotopes that have a measured $\bar{\nu}$ in the data library
will emit fission gammas or neutrons. At present the gammas and
neutrons are emitted isotropically. The data and empirical models are
described in detail in the following subsections.

\input{physics}

\subsection{Implementation}

\input{implementation}

Different sampling methods can be selected by calling the following functions.
\input{arg}

\clearpage
\begin{thebibliography}{99}
\input{fission_geant_bib}
\end{thebibliography}

\end{document}
